---
title: "Exercise 1, Keel Scruton"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages and import data

```{r}
#install.packages("readr")
library(tidyverse)
data <- read_csv('/Users/keelscruton/Desktop/Org Network Analysis/2022-ona-assignments/Exercises/ex1/Connections.csv')
```

Here we want to obtain the counts of by employer and total count as well. 

```{r}
#Counts by employer 
library(dplyr)
data %>% count(Company) %>% arrange(-n)
#Total count is the length of the data frame (data), can also be counted using count function. 
count(data)

```

Next we create nodes and edges data frames, these will be used with igraph. 

The first step will be to create a new data set that is the unique contacts (first name + letter of first name) this is the node list (or dataframe)

```{r}

#Make a column with the first name and first letter of the last name (first_last as variable name)

data$first_last <- paste(data$`First Name`,substr(data$`Last Name`,1,1))
```

```{r}
#create a distinct list of all contact first_last names, name this column label. 
nodes <- data %>%
  distinct(`first_last`) %>%
  rename(label = first_last)

#set a unique ID value. 
nodes <- nodes %>% rowid_to_column("id")

```

The next step is to create the Edge List
```{r}

edges =data %>% select(first_last, Company) %>%
  inner_join(., select(., first_last, Company), by = "Company") %>%
  rename(INDIVIDUAL1 = first_last.x, INDIVIDUAL2 = first_last.y) %>%
  filter(INDIVIDUAL1 != INDIVIDUAL2) %>%
  unique #%>% select(-Company)

#want to replace each name with the unique id's made

edges2 <- edges %>% 
  left_join(nodes, by = c("INDIVIDUAL1" = "label")) %>% 
  rename(person1 = id)

edges2 <- edges2 %>% 
  left_join(nodes, by = c("INDIVIDUAL2" = "label")) %>% 
  rename(person2 = id)

edges_final <- select(edges2,person1,person2)

```


Now we can plot the network 

```{r}
library(tidygraph)
library(ggraph)

routes_tidy <- tbl_graph(nodes = nodes, edges = edges_final, directed = FALSE)

routes_tidy %>% 
  activate(edges) 


ggraph(routes_tidy) + geom_edge_link() + geom_node_point() + theme_graph()
```



