---
title: "Exercise 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#make sure to load the proper packages that are required
library("tidyverse")
library("igraph")
library('ggraph')
library("tidygraph")
```

The node and edge lists were made in excel files to be able to import them easily. 

```{r}
nodes <- read_csv('nodes_list.csv')
edges <- read_csv('edge_list.csv')

```

Next we can create the potential relationship network. this is used to determine which seat has the most potential. 

```{r}

relationship_network <- tbl_graph(nodes=nodes, edges=edges, directed=FALSE)
Degree <- degree(relationship_network)
print(Degree)

closeness <- closeness(relationship_network)
print(closeness)


```

For evaluating which seat is best I will take into account degree and closeness.  Should note that the only seats available are ABCD (the last 4 in each list printed above) based on degree B,C and D all have the same degree. Based on the closeness which is how central or close a node is to all other nodes B and C have the highest values and are equal (more desirable centrality). Based on this analysis seats B or C are equivalently connected to the other seats on the bus. 


Should note that if you select B you risk that D, C and A are not filled, if you select C you also risk that D B and A are not filled so each option has equal risk and as such either of them can be selected. 


Now we can plot the network. 
```{r}
plot.igraph(relationship_network, 
            vertex.size=degree(relationship_network), 
            main="Degree")
```




