---
title: "Exercise_3_ONA_Keel_Scruton"
output: pdf_document
---

Import/load all required packages. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(arrow)
library(ggraph)
library(igraph)
```


## Load data

Load the following data:
  + applications from `app_data_sample.parquet`
  + edges from `edges_sample.csv`

```{r load-data}
data_path <- "/Users/keelscruton/Desktop/Org Network Analysis/672_project_data/"
applications <- read_parquet(paste0(data_path,"app_data_sample.parquet"))
edges <- read_csv(paste0(data_path,"edges_sample.csv"))
applications
edges
```


## Get gender for examiners

We'll get gender based on the first name of the examiner, which is recorded in the field `examiner_name_first`. We'll use library `gender` for that, relying on a modified version of their own [example](https://cran.r-project.org/web/packages/gender/vignettes/predicting-gender.hedges_examinerl).

Note that there are over 2 million records in the applications table -- that's because there are many records for each examiner, as many as the number of applications that examiner worked on during this time frame. Our first step therefore is to get all *unique* names in a separate list `examiner_names`. We will then guess gender for each one and will join this table back to the original dataset. So, let's get names without repetition:

```{r gender-1}
library(gender)
#install_genderdata_package() # only run this line the first time you use the package, to get data for it
# get a list of first names without repetitions
examiner_names <- applications %>% 
  distinct(examiner_name_first)
examiner_names
```


Now let's use function `gender()` as shown in the example for the package to attach a gender and probability to each name and put the results into the table `examiner_names_gender`

```{r gender-2}
# get a table of names and gender
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )
examiner_names_gender
```


Finally, let's join that table back to our original applications data and discard the temporary tables we have just created to reduce clutter in our environment.

```{r gender-3}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)
# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")
# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```


## Guess the examiner's race

We'll now use package `wru` to estimate likely race of an examiner. Just like with gender, we'll get a list of unique names first, only now we are using surnames.

```{r race-1}
library(wru)
examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()
examiner_surnames
```
We'll follow the instructions for the package outlined here [https://github.com/kosukeimai/wru](https://github.com/kosukeimai/wru).

```{r race-2}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()
examiner_race
```

As you can see, we get probabilities across five broad US Census categories: white, black, Hispanic, Asian and other. (Some of you may correctly point out that Hispanic is not a race category in the US Census, but these are the limitations of this package.)



Our final step here is to pick the race category that has the highest probability for each last name and then join the table back to the main applications table. See this example for comparing values across columns: [https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/](https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/). And this one for `case_when()` function: [https://dplyr.tidyverse.org/reference/case_when.hedges_examinerl](https://dplyr.tidyverse.org/reference/case_when.hedges_examinerl).

```{r race-3}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))
examiner_race
```

Let's join the data back to the applications table.

```{r race-4}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)
applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))
rm(examiner_race)
rm(examiner_surnames)
gc()
```

## Examiner's tenure 

To figure out the timespan for which we observe each examiner in the applications data, let's find the first and the last observed date for each examiner. We'll first get examiner IDs and application dates in a separate table, for ease of manipulation. We'll keep examiner ID (the field `examiner_id`), and earliest and latest dates for each application (`filing_date` and `appl_status_date` respectively). We'll use functions in package `lubridate` to work with date and time values.

```{r tenure-1}
library(lubridate) # to work with dates
examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 
examiner_dates
```
The dates look inconsistent in terms of formatting. Let's make them consistent. We'll create new variables `start_date` and `end_date`.

```{r tenure-2}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

Let's now identify the earliest and the latest date for each examiner and calculate the difference in days, which is their tenure in the organization.

```{r tenure-3}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)
examiner_dates
```

Joining back to the applications data.

```{r tenure-4}
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")
rm(examiner_dates)
gc()
```

##Begining of work not included in class github. 
### Section 2: 
The two work groups (groups) we will focus on are 161 and 162
Here we create subgroups that only contain those instances. We will also ensure we only
display unique instances of each examiner. 
```{r}
vars <- c("gender","race","tenure_days")
applications = drop_na(applications,any_of(vars))

```

```{r}


group_161 <- applications[substr(applications$examiner_art_unit, 1, 3) == 161,]
group_161 <-group_161[row.names(unique(group_161[,"examiner_id"])),]


group_162 <- applications[substr(applications$examiner_art_unit, 1, 3) == 162,]
group_162 <-group_162[row.names(unique(group_162[,"examiner_id"])),]

```

create some summary statistics 

```{r}
summary(group_161)
print('----------------------Next Group ---------------------------')
summary(group_162)
```

Next we can create some plots to better compare the two groups based on demographics. 

##Gender breakdown
First we will compare based on gender. 
```{r}
library(gridExtra)
gender_161 <- ggplot(data=group_161, aes(x=gender)) +
  geom_bar(aes(y = (..count..)/sum(..count..)) )  +
  ylab("Proportion")+
  xlab("Gender")+
  ylim(0,1)+
  ggtitle(paste0("Gender Comparison Group 161"))



gender_162 <- ggplot(data=group_162, aes(x=gender)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  ylab("Proportion")+
  xlab("Gender")+
  ylim(0,1)+
  ggtitle(paste0("Gender Comparison Group 162"))

grid.arrange(gender_161,gender_162,ncol=2, widths=c(1,1))

```
From the above plots we can see that group 161 has more male examiners, group 162 has more female examiners. Should note that all NA values for gender were dropped, these values could not be estimated based on our methods and so in reality could alter the results shown here. we can possibly assume that the distribution would be representative of the distributions we see here. 



##Race breakdown
Next we will compare based on race 
```{r}
library(gridExtra)
race_161 <- ggplot(data=group_161, aes(x=race)) +
  geom_bar(aes(y = (..count..)/sum(..count..)) )  +
  ylab("Proportion")+
  xlab("Race")+
  ylim(0,1)+
  ggtitle(paste0("Race Comparison Group 161"))



race_162 <- ggplot(data=group_162, aes(x=race)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  ylab("Proportion")+
  xlab("Race")+
  ylim(0,1)+
  ggtitle(paste0("Race Comparison Group 162"))

grid.arrange(race_161,race_162,ncol=2, widths=c(1,1))
```
From these results we can see that group 161 and 162 both have more white examiners than any other group. at just over 75% for each, as well both groups also have just under 25% of the examiner population being Asian. the biggest difference is that there is slightly more black examiners in group 161 then in 162 and there apears to be some minimal number of hispanic examiners in group 162 that is not present in 161. 



##Tenure breakdown
Next we will compare based on tenure  (tenure days)
```{r}
library(gridExtra)
a= ggplot(data=group_161, aes(group_161$tenure_days)) + geom_histogram()+  ggtitle(paste0("Tenure Comparison Group 161"))+ylab("Count")+
  xlab("Tenure Days")

b= ggplot(data=group_162, aes(group_162$tenure_days)) + geom_histogram() +ggtitle(paste0("Tenure Comparison Group 162"))+ylab("Count")+
  xlab("Tenure Days")




grid.arrange(a,b,ncol=2, widths=c(1,1))




```

Observing the tenure differences between the groups we can see that the tenure distributions are very similar above the 6000 days mark (~16 years). However in group 161 there is a group of examiners who have a shorter tenure under 5500 days. Both groups are mostly made up of examiners who have been there at least 5000 days (~13 years)


### Section 3:

Create advice networks from `edges_sample` and calculate centrality scores for examiners in your selected workgroups
```{r}
#create distinct subset of examiners with only the art unit and examiner id to be able to re join onto our other data afterwards
examiner_dis = distinct(subset(applications, select=c(examiner_art_unit, examiner_id)))
examiner_dis$group = substr(examiner_dis$examiner_art_unit, 1,3)
#get rid of all examiners except those in group 161 or 162
examiner_dis = examiner_dis[examiner_dis$group==161 | examiner_dis$group==162,]

```

Now that we have a list of the examiners who are part of work groups 161 and 162 we can combine (merge) it with the edge list (edges) this will allow us to form our subset network. 

```{r}
edges_examiner = merge(x=edges, y=examiner_dis, by.x="ego_examiner_id", by.y="examiner_id", all.x=TRUE)
edges_examiner = edges_examiner %>% rename(ego_au=examiner_art_unit, ego_group=group)

edges_examiner = merge(x=edges_examiner, y=examiner_dis, by.x="alter_examiner_id", by.y="examiner_id", all.x=TRUE)
edges_examiner = edges_examiner %>% rename(alter_au=examiner_art_unit, alter_group=group)

edges_examiner = drop_na(edges_examiner) #drop all na, ie values not in the selected workgroups. 

```

Now the above data set has the edges of alter and ego examiners. we can next create the list of nodes for both the ego and alter examiners. 

```{r}
Ego_nodes = subset(edges_examiner, select=c(ego_examiner_id,ego_au, ego_group)) %>% rename(examiner_id=ego_examiner_id,art_unit=ego_au,group=ego_group)

Alter_nodes = subset(edges_examiner, select=c(alter_examiner_id,alter_au, alter_group))%>% rename(examiner_id=alter_examiner_id,art_unit=alter_au,group=alter_group)

Nodes_full = distinct(rbind(Ego_nodes, Alter_nodes))

#Nodes_full is a list of all the distinct examiners involved in the advice network

```
Now we can create the graph network. 
```{r}
#ran into error where i had duplicated vertex names, to fix take the first instances only in the nodes list. 

Nodes_full = Nodes_full %>% group_by(examiner_id) %>% summarise(examiner_id=first(examiner_id), art_unit=first(art_unit), group=first(group))

network <- graph_from_data_frame(d=edges_examiner, vertices=Nodes_full, directed=TRUE)


#Print the simple network with no account for degree or visualisation of the two units
ggraph(network, layout="kk") +
  geom_edge_link()+
  geom_node_point()

```

To be able to make a better graph we need to calculate the degree of each node. as well 
we need to calculate a measure of centrality. As this is an advice network, i am assuming that it would be the most advantageous to be able to reach out easily across the network to the most people possible to be able to get help with a patent case. as such I believe that closeness is the most advantageous. This value would help find the examiners who are best connected to the network for efficiency of informational travel between nodes. 
```{r}
Degree <- degree(network)
Closeness <- closeness(network)

ggraph(network, layout="kk") +
  geom_edge_link(arrow = arrow(length = unit(1, 'mm')), 
                 end_cap = circle(1.2, 'mm'))+
  geom_node_point(aes(size=Degree, color=Nodes_full$group), show.legend=T)


```

From the above graph we can see that those in group 161 have much bigger nodes, meaning that each examiner seems to be connected to more people than those examiners in group 162. As well in the bottom right we can see there seems to be a segmented group of connected examiners from group 162 whereas the group of connected examiners from 161 is also fairly well connected to other members from 162. we also have a few outliers of small connections that do not mix or connect with the main group and are isolated. Just visually observing the arrows that represent direction, there seem to be many examiners in the 161 group that have many connections (degree) but have arrows going away from them, (they seek lots of advice?) There are also equaly a few nodes in the 162 clusters that everyone seems to come to for advice. 

Contrasting with the results of the demographic analysis we found that group 161 has more males and has a greater distribution across tenure. The greater distribution of tenure could explain why those examiners are more willing/needing of seeking out advice, and requiring a strong network in order to get the work done effectively. Effectively there was no real differences between the two groups on race so it is probably not having an effect here. group 162 that had more female examiners seem to go to themselves for advice more (members of the same group) but i doubt this is correlated because the proportion of male to female in both groups is not very different (very small difference in proportion)

```{r}
ggraph(network, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=Closeness, color=Nodes_full$group), show.legend=T)



```





